---
title: "Lab 09 - Logistic Regression"
author: Chu-Chun Ku
output:
  html_document:
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package and suppress scientific notation 
```{r}
# load packages 
library(caret) 
library(gains)
options(scipen=999)

ebay.df <- read.csv("eBayAuctions.csv", stringsAsFactors = FALSE)
# first six rows
head(ebay.df)
names(ebay.df)
# not display scientific notation

```

## 1 create a data frame
```{r}
# load the data
ebay.df <- read.csv("eBayAuctions.csv", stringsAsFactors = FALSE)

# variable names 
head(ebay.df)
names(ebay.df)
```

## 2 apply a function to each group 
```{r}
# mean of the binary outcome for each day of the week of auction close   
eachday <- tapply(ebay.df$Competitive, ebay.df$endDay, mean)
# mean of the binary outcome for each auction category  
eachcategory <- tapply(ebay.df$Competitive, ebay.df$Category, mean)
```

## 3 reduce levels in the day of week that the auction closed  
```{r}
# combine Sunday and Friday into a single category called Sun_Fri
ebay.df$endDay[ebay.df$endDay == "Sun"] <- "Sun_Fri"
ebay.df$endDay[ebay.df$endDay == "Fri"] <- "Sun_Fri"

```

## 4 reduce levels in auction categories 
```{r}
# combine Business/Industrial and Computers into a single category called Computers
ebay.df$Category[ebay.df$Category == "Business/Industrial"] <- "Computers"

# combine Antique/Art/Craft and Collectibles into a single category called collectibles
ebay.df$Category[ebay.df$Category == "Antique/Art/Craft"] <- "Collectibles"
```

## 5 convert Duration to a categorical or factor variable 
```{r}
# convert Duration from a numeric variable to a categorical or factor variable  
ebay.df$Duration <- as.factor(ebay.df$Duration)
# variable types 
str(ebay.df)
```

## 6 data partition
```{r}
# set seed for reproducing the partition 
set.seed(1)
# takes a sample of row numbers for the training set 
a <- dim(ebay.df)[1]
train.index <- sample(1:a, size = 0.6 * a)
# training set 
train.df <- ebay.df[train.index, ]
# test set 
test.df <- ebay.df[-train.index, ]
```

## 7 fit a logistic regression model 
```{r,warning=FALSE}
# logistic regression using all the variables
reg <- glm(Competitive. ~ ., data = train.df, family = "binomial")
# summary table 
summary(reg)
```

## 8 generate predicted probabilities for records in the test set 
```{r}
# predicted probabilities 
pred <- predict(reg, newdata = test.df, type = "response")
# first six values 
head(pred)
```

## 9 confusion matrix 
```{r}

cm1 <- confusionMatrix(as.factor(ifelse(pred > 0.5, 1, 0)),
                       as.factor(test.df$Competitive.),
                       positive = "1")
as.table(cm1)
```

## 10 creating a gain table [optional]
```{r}
# gain table 
str(test.df$Competitive)
# cumulative percentage of competitive auctions 

# cumulative number of auctions 

```

## 11 plot a lift chart [optional]
```{r}
# plot the cumulative number of competitive auctions against the cumulative number of auctions

# add a baseline curve 

```