---
title: "Lab 15 - Neural networks"
author: Chu-Chun Ku
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
knitr::knit_meta(clean=T)
```

## 0 load the packages 
```{r}
library(neuralnet) 
library(caret)

```

## 1 create a data frame
```{r}
# load the data 
df <- read.csv("EastWestAirlinesNN.csv")
# first six rows 
head(df)
# column names
names(df)
```

## 2 create two outcome dummies to represent output nodes  
```{r}
# create a variable named purchase
df$purchase <- df$Phone_sale == 1
# create a variable named not_purchase
df$not_purchase <- df$Phone_sale == 0
```

## 3 remove unnecessary variables
```{r}
# remove ID column and update df 
df <- df[, -1]
# column names 
names(df)
```

## 4 remove rows with missing values
```{r}

df <- na.omit(df)
```

## 5 data partition 
```{r}
# set the seed 
set.seed(1)
# row indexes of the training set
train.index <- sample(c(1:dim(df)[1]) , 0.6 * dim(df)[1])
# training set
train.df <- df[train.index, ]
# test set
test.df <- df[-train.index, ]
```

## 6 normalize predictors to a scale of [0, 1] 
```{r}
# estimate the transformation 
norm.values <- preProcess(train.df[,c(1:14)], method="range") 
# normalize the numerical predictors in the training set 
train.norm.df <- predict(norm.values, train.df[,c(1:14)])
# normalize the numerical predictors in the test set 
test.norm.df <- predict(norm.values, test.df[,c(1:14)])
```

## 7 fit a neural network on the training set 
```{r}
# set the random seed 
set.seed(1)
# fit a neural network that has a single hidden layer with 5 nodes (should take ~ 1 minute)
nn <- neuralnet(train.df$not_purchase + train.df$purchase ~ Topflight + Balance + Qual_miles + cc1_miles. + cc2_miles. + cc3_miles. + Bonus_miles + Bonus_trans + Flight_miles_12mo + Flight_trans_12 + Online_12 + Email + Club_member + Any_cc_miles_12mo, data=train.norm.df, linear.output = FALSE, hidden = 5)
# plot the network
plot(nn, rep = "best")
# display weights
nn$weights
```

## 8 generate predicted probabilities for records in the test set  
```{r}
# make predictions using compute() 
nn.pred <- compute(nn, test.norm.df)

# predicted probabilities of making purchases
nn.pred.prob <- nn.pred$net.result[, 2] / (nn.pred$net.result[, 1] + nn.pred$net.result[, 2])

```

## 9 generate predicted classes for records in the test set 
```{r}
nn.pred.class <- ifelse(nn.pred.prob > 0.5, 1, 0)

```

## 10 create a confusion matrix for records in the test set
```{r}
confusionMatrix(as.factor(nn.pred.class), as.factor(test.df$Phone_sale), positive = "1")


```
