---
title: "Lab 03 - Data Wrangling"
author: Chu-Chun Ku
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: simplex
    number_sections: false
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## 0 load the package 
```{r}
library(tidyverse)
```

## 1 create a tibble 

```{r}
# load the data 
df <- read_csv("firm.csv")
# first six rows 
head(df)
# number of rows  
nrow(df)
# number of columns 
ncol(df)
# column names 
names(df)
```

## 2 subset data 

### 2.1 one column 
```{r}
df$NI
select(df,NI)
```

### 2.2 three columns 
```{r}
df[,c(3,4,8)]
select(df, FYEAR, TIC, NI)
```

### 2.3 four rows and three columns
```{r}
df[c(1:4),c(3,4,8)]

```

## 3 filter rows  

### 3.1 observations that belong to AAPL in 2015 
```{r}
#Pick rows that TIC is equal to AAPL and FYEAR is equal to 2015 using df[,] and filter()
df[df$TIC=="AAPL" & df$FYEAR==2015, ]
filter(df,TIC=="AAPL", FYEAR==2015 )
```

### 3.2 observations that have the minimum NI value 
```{r}
#Pick rows that NI is equal to the minimum NI value using df[,] and filter().
min(df$NI)
df[df$NI==min(df$NI),]
filter(df, df$NI==min(df$NI))
```

### 3.3 select columns of observations that belong to AAPL in 2015   

```{r}
#Select FYEAR, TIC, NI columns and pick rows that TIC is equal to AAPL and FYEAR is equal to 2015 using df[,].
df[df$TIC=="AAPL" & df$FYEAR==2015, c("FYEAR","TIC","NI") ]
```

## 4 arrange rows 

### 4.1 ascending order of NI
```{r}
# Reorder df in an ascending order of NI using arrange().
arrange(df,NI)
```

### 4.2 descending order of NI
```{r}
# Reorder df in a descending order of NI using arrange() with desc()
arrange(df, desc(NI))
```

### 4.3 ascending order of TIC and FYEAR, and descending order of NI

```{r}
# Reorder df in an ascending order of TIC and FYEAR, and descending order of NI using arrange()
arrange(df, TIC, FYEAR, desc(NI))

```

## 5 create new variables  

### 5.1 return on assets 

```{r}
# Create a new column named ROA in df, which divides net income (NI) by total assets (AT).
df$ROA <- df$NI / df$AT
head(df)
```

### 5.2 net income bin numbers  
```{r}
# step1: identify thresholds of 4 bins with an equal number of observations 
bins <- quantile(df$NI, probs = c(0,0.25,0.5,0.75,1))

# step2: remove the names of a named vector  
bins <- unname(bins)

# step3: bin NI  
df$NI_bin <- .bincode(df$NI, bins, include.lowest = TRUE)
table(df$NI_bin)

```

### 5.3 log of assets 
```{r}
# Create a new column named AT_LOG in df which computes the log value of total assets (AT) using mutate(). Use head() to print the first six rows of df.
df <- mutate(df, AT_LOG = log(df$AT))
head(df)
```

### 5.4 drop a column 

```{r}
# Remove return on assets (ROA) from df by assigning NULL to the column. Use head() to print the first six rows of df.
df$ROA <- NULL
head(df)

```

